---
title: "Cross Correlation"
output:
  html_document:
    df_print: paged
---

```{r Package Loading,include=FALSE}
library(tidyverse)
library(lubridate)
library(plotly)
library(gridExtra)
library(roll)
library(plotly)
library(kableExtra)

# This script is for cross correlation of rainfall and water level data
setwd("E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/Plots/CrossCorr")
```

# 01 Plotting Raw and Rolling Mean data

Some testing was done on length of rolling mean. I applied 6 hrs (24 15min timestamps), 12 hrs (48), and 24 (96). Since early tests of cross correlation indicated that rain preceded water level change by >1 day typically, I have carried out the cross correlation here with 12 hr rolling mean. These top interactive plots just show the difference between the raw and rolling mean data for water level. The final plot is the 12 hr rolling mean (in blue) and raw (faint grey..) rainfall data from Ancient Forest Station.


```{r,echo=FALSE, message=FALSE,warning=FALSE}
rain <- read_csv("E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/AncientForest2019Tidied.csv") %>%
  dplyr::select(timestamp, rain_mm_tot) %>% 
  dplyr::mutate(mn.rain.24 = roll_mean(rain_mm_tot, width = 24),
                mn.rain.48 = roll_mean(rain_mm_tot, width = 48),
                mn.rain.96 = roll_mean(rain_mm_tot, width = 96))


files <- dir(pattern = "*-G.csv", path = "E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/") # pull the file names from folder

#files <- "A1-G.csv"


# Correlate and Plot Function ---------------------------------------------


WlBinder <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no) %>% 
    dplyr::mutate(mn.val.24 = roll_mean(raw_val, width = 24),
                  mn.val.48 = roll_mean(raw_val, width = 48),
                  mn.val.96 = roll_mean(raw_val, width = 96))
  
  
  # Join the data together and drop all timestamps with NA -----------------
  
  dat <- full_join(rain,wl) %>% 
    drop_na() %>% 
    distinct() %>% 
    filter(date(timestamp) >= "2019-04-01" & date(timestamp) <= "2019-11-01")
  
}

data <- files %>% 
  map(WlBinder) %>% 
  reduce(rbind)

# p24 <- data %>% 
#   ggplot()+
#   geom_line(aes(timestamp, raw_val, colour = site.no), size = 1)+
#   geom_line(aes(timestamp, mn.val.24, group = site.no), linetype="dashed")+
#   labs(title = "6Hr rolling avg")
#   
# ggplotly(p24)

p48 <- data %>% 
  ggplot()+
  geom_line(aes(timestamp, raw_val, colour = site.no), size = 1)+
  geom_line(aes(timestamp, mn.val.48, group = site.no), linetype="dashed")+
  labs(title = "12Hr rolling avg")
  
p48

rain.p48 <- data %>% 
  ggplot()+
  geom_col(aes(timestamp,rain_mm_tot))+
  geom_col(aes(timestamp,mn.rain.48),colour = "#6699CC75")+
  ylim(c(0,10))+
  labs(title = "12 hr rolling mean rainfall (blue)")
rain.p48


```



# 02 Cross Correlation Plots over Full Fimespan

## Raw and 12 hr rolling mean cross correlation from entire dataset

When I correlate the raw & 12 hr rolling mean datasets over the full period, the influence of snowmelt decreases teh significance of the cross correlation between rain and water levels. The rolling mean does appear to improve correlation coefficients but the water level data clearly have trends that reduce these.

```{r Raw data cross corr,echo=FALSE, message=FALSE,warning=FALSE}
rain <- read_csv("E:/Jeremy's MSc Research/Hydrometric and GIS/Climate Data/AncientForest2019Tidied.csv") %>%
  dplyr::select(timestamp, rain_mm_tot)


files <- dir(pattern = "*-G.csv", path = "E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/") # pull the file names from folder

#files <- "A1-G.csv"


# Correlate and Plot Function ---------------------------------------------


CrossCorrelateR <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no)
  
  
  # Join the data together and drop all timestamps with NA -----------------
  
  dat <- full_join(rain,wl) %>% 
    drop_na() %>% 
    distinct() %>% 
    filter(date(timestamp) >= "2019-04-01" & date(timestamp) <= "2019-11-01")
  
  bccf <-  ccf(dat$rain_mm_tot, dat$raw_val,plot = FALSE, lag = 250)
  
  bacfdf <- with(bccf, data.frame(lag, acf))%>%
    rename(crcor = "acf", lagg = "lag") %>% 
    mutate(site = sites)

}

data <- files %>% 
  map(CrossCorrelateR) %>% 
  reduce(rbind)

A <- data %>% filter(site == "A2-G")


conf.level <- 0.95
ciline <- qnorm((1 - conf.level)/2)/sqrt(length(A$lagg))

p <- data %>% 
  ggplot()+
  geom_line(aes(lagg,crcor,colour = site), size = .9)+ 
  geom_hline(aes(yintercept = 0)) +
  geom_hline(aes(yintercept = ciline), linetype = 2, color = 'darkblue') +
  geom_hline(aes(yintercept = -ciline), linetype = 2, color = 'darkblue')+
  # geom_segment(mapping = aes(xend = lagg, yend = 0))+
  labs(x = "Lag (15 Min)", y = "Cross-Correlation", title = paste("Raw Data Rain to Water Level Correlation"))
  
p 
```

```{r 12hr roll mean cross corr,echo=FALSE, message=FALSE, warning=FALSE}
rain <- read_csv("E:/Jeremy's MSc Research/Hydrometric and GIS/Climate Data/AncientForest2019Tidied.csv") %>%
  dplyr::select(timestamp, rain_mm_tot) %>% 
  dplyr::mutate(mn.rain = roll_mean(rain_mm_tot, width = 48))


files <- dir(pattern = "*-G.csv", path = "E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/") # pull the file names from folder

#files <- "A1-G.csv"


# Correlate and Plot Function ---------------------------------------------


CrossCorrelateR <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no) %>% 
    dplyr::mutate(mn.val = roll_mean(raw_val, width = 48))
  
  
  # Join the data together and drop all timestamps with NA -----------------
  
  dat <- full_join(rain,wl) %>% 
    drop_na() %>% 
    distinct() %>% 
    filter(date(timestamp) >= "2019-04-01" & date(timestamp) <= "2019-11-01")
  
  bccf <-  ccf(dat$mn.rain, dat$mn.val,plot = FALSE, lag = 250)
  
  bacfdf <- with(bccf, data.frame(lag, acf))%>%
    rename(crcor = "acf", lagg = "lag") %>% 
    mutate(site = sites)

}

data <- files %>% 
  map(CrossCorrelateR) %>% 
  reduce(rbind)

A <- data %>% filter(site == "A2-G")


conf.level <- 0.95
ciline <- qnorm((1 - conf.level)/2)/sqrt(length(A$lagg))

p <- data %>% 
  ggplot()+
  geom_line(aes(lagg,crcor,colour = site), size = .9)+ 
  geom_hline(aes(yintercept = 0)) +
  geom_hline(aes(yintercept = ciline), linetype = 2, color = 'darkblue') +
  geom_hline(aes(yintercept = -ciline), linetype = 2, color = 'darkblue')+
  # geom_segment(mapping = aes(xend = lagg, yend = 0))+
  labs(x = "Lag (15 Min)", y = "Cross-Correlation", title = paste("12 hr rollmean Rain to Water Level Correlation"))
  
p 
```

# 03 Cross Correlation of Rain Only Period
Trimming the analysis period to the truly snow free period improves correlation significance in the raw data, with even more improved 12 hr mean correlation. This suggests that the snowmelt influence does affect this analysis strongly, as does noise in the water level data.The rolling mean also generalizes the rain data to reduce the impact of sporadic rain versus periods of rain.


```{r Raw data from 01 June to 01 Nov,echo=FALSE, message=FALSE,warning=FALSE}
rain <- read_csv("E:/Jeremy's MSc Research/Hydrometric and GIS/Climate Data/AncientForest2019Tidied.csv") %>%
  dplyr::select(timestamp, rain_mm_tot)


files <- dir(pattern = "*-G.csv", path = "E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/") # pull the file names from folder

#files <- "A1-G.csv"


# Correlate and Plot Function ---------------------------------------------


CrossCorrelateR <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no)
  
  
  # Join the data together and drop all timestamps with NA -----------------
  
  dat <- full_join(rain,wl) %>% 
    drop_na() %>% 
    distinct() %>% 
    filter(date(timestamp) >= "2019-06-01" & date(timestamp) <= "2019-11-01")
  
  bccf <-  ccf(dat$rain_mm_tot, dat$raw_val,plot = FALSE, lag = 250)
  
  bacfdf <- with(bccf, data.frame(lag, acf))%>%
    rename(crcor = "acf", lagg = "lag") %>% 
    mutate(site = sites)

}

data <- files %>% 
  map(CrossCorrelateR) %>% 
  reduce(rbind)

A <- data %>% filter(site == "A2-G")


conf.level <- 0.95
ciline <- qnorm((1 - conf.level)/2)/sqrt(length(A$lagg))

p <- data %>% 
  ggplot()+
  geom_line(aes(lagg,crcor,colour = site), size = .9)+ 
  geom_hline(aes(yintercept = 0)) +
  geom_hline(aes(yintercept = ciline), linetype = 2, color = 'darkblue') +
  geom_hline(aes(yintercept = -ciline), linetype = 2, color = 'darkblue')+
  # geom_segment(mapping = aes(xend = lagg, yend = 0))+
  labs(x = "Lag (15 Min)", y = "Cross-Correlation", title = paste("Raw Rain to Water Level Correlation 01 Jun - 01 Nov 2019"))
  
p 
```


```{r 12 hr rolling mean, 01 June to 01 Nov,echo=FALSE, message=FALSE, warning=FALSE}
rain <- read_csv("E:/Jeremy's MSc Research/Hydrometric and GIS/Climate Data/AncientForest2019Tidied.csv") %>%
  dplyr::select(timestamp, rain_mm_tot) %>% 
  dplyr::mutate(mn.rain = roll_mean(rain_mm_tot, width = 48))


files <- dir(pattern = "*-G.csv", path = "E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/") # pull the file names from folder

#files <- "A1-G.csv"


# Correlate and Plot Function ---------------------------------------------


CrossCorrelateR <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("E:/Jeremy's MSc Research/Hydrometric and GIS/Water Data/Groundwater/For Analysis/CrossCorr/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no) %>% 
    dplyr::mutate(mn.val = roll_mean(raw_val, width = 48))
  
  
  # Join the data together and drop all timestamps with NA -----------------
  
  dat <- full_join(rain,wl) %>% 
    drop_na() %>% 
    distinct() %>% 
    filter(date(timestamp) >= "2019-06-01" & date(timestamp) <= "2019-11-01")
  
  bccf <-  ccf(dat$mn.rain, dat$mn.val,plot = FALSE, lag = 250)
  
  bacfdf <- with(bccf, data.frame(lag, acf))%>%
    rename(crcor = "acf", lagg = "lag") %>% 
    mutate(site = sites)

}

data <- files %>% 
  map(CrossCorrelateR) %>% 
  reduce(rbind)

A <- data %>% filter(site == "A2-G")


conf.level <- 0.95
ciline <- qnorm((1 - conf.level)/2)/sqrt(length(A$lagg))

p <- data %>% 
  ggplot()+
  geom_line(aes(lagg,crcor,colour = site), size = .9)+ 
  geom_hline(aes(yintercept = 0)) +
  geom_hline(aes(yintercept = ciline), linetype = 2, color = 'darkblue') +
  geom_hline(aes(yintercept = -ciline), linetype = 2, color = 'darkblue')+
  # geom_segment(mapping = aes(xend = lagg, yend = 0))+
  labs(x = "Lag (15 Min)", y = "Cross-Correlation", title = paste(" 12 hr rollmean Rain to Water Level Correlation 01 Jun - 01 Nov 2019"))
  
p 


```


#  04 Summary Values
## Peak positive correlation values over 95% confidence (From last plot)
Wells A1-G, B4-G, and C2-G did not have significant correlations over the 2019 period. A1-G water table dynamics are likely different from all other sites due to the site being located in peat as opposed to sediment. The timeseries from B4-G is extremely spiky which appears most likely to be due to sensor noise, though could also be due to an open well head causing artificially rapid recharge. The C transect wells are higher in the flow path, in coarser sediments and within an area with limited ponding. It is not clear why the C2-G correlation does not exceed the 95% CI, however its correlation matches well with others in that transect.

```{r,echo=FALSE, message=FALSE, warning=FALSE}
lag.dat <- data %>%
  dplyr::group_by(site) %>% 
  dplyr::filter(crcor == max(crcor) & crcor >= 0.08756473) %>% 
  dplyr::mutate(lag.hrs = lagg/4) %>% 
  dplyr::select(site,lagg,lag.hrs,crcor)

kable(lag.dat) %>% 
  kable_material()
```


## Mean and standard deviation of significant correlations
The time lag between rainfall and water level timeseries correlation is an average of 20.6 hrs, with a standard deviation of 15.6 hrs. This range may be due to a range in infiltration rates (due to range in sediment composition), variability in seal of well head at the ground surface (creating artificially fast recharge pathways, likely at A4-G) or ponding at sites lower in the gradient (rain direct to an above ground water table).

```{r,echo=FALSE, message=FALSE, warning=FALSE}
lag.summary <- lag.dat %>%
  dplyr::ungroup() %>% 
  dplyr::summarise(mn.lag.hrs = mean(lag.hrs),
                   sd.lag.hrs = sd(lag.hrs))

kable(lag.summary) %>% 
  kable_material()
```


