
```{r Chunk Settings, include=FALSE}
knitr::opts_chunk$set(message = FALSE,warning = FALSE, echo = FALSE)
```


```{r Packages}
library(tidyverse)
library(lubridate)
library(plotly)
library(RcppRoll)
library(broom)
```


it will be best to avg on daily timescale before bonding with tipping bucket data.


# Read in & Tidy Regional Data

Read in the regional precip dat (Longworth & Dome Mtn) This is cumulative precip data (from a pluvio?), join together, fix col headers, add Rain_Tot data col (not cumulative)

```{r Read and Tidy Data}
longprecip <- read_csv("Precipitation-Gap-Filling/LongworthCumPrecip.csv", skip = 2) %>% 
  mutate(Site = "Longworth")

domeprecip <- read_csv("Precipitation-Gap-Filling/DomeMtnCumPrecip.csv", skip = 2) %>% 
  mutate(Site = "DomeMtn")

precipdat <- full_join(longprecip,domeprecip)

colnames(precipdat) <- c("Timestamp", "CuPrecip", "Grade", "Approval","Interpolation", "event", "Site")

precipdat <- precipdat%>%
  mutate(date = date(Timestamp)) %>% 
  group_by(Site, date) %>% 
  summarise(CuPrecip = mean(CuPrecip, na.rm = T)) %>% 
  mutate(Rain_Tot = CuPrecip - lag(CuPrecip, 1))

precipdat$Rain_Tot[precipdat$Rain_Tot < 0] = 0

precipdat <- precipdat %>% 
  pivot_longer(cols = c(Rain_Tot,CuPrecip), names_to = "Parameter")

# Read in the regional Temperature dat (Longworth & Dome Mtn) Units are celsius, join together, fix col headers.

longairt <- read_csv("Precipitation-Gap-Filling/LongworthAirT.csv", skip = 2) %>% 
  mutate(Site = "Longworth") 

domeairt <- read_csv("Precipitation-Gap-Filling/DomeMtnAirT.csv", skip = 2) %>% 
  mutate(Site = "DomeMtn")

tempdat <- full_join(longairt,domeairt)

colnames(tempdat) <- c("Timestamp", "AirTC", "Grade", "Approval","Interpolation", "event", "Site")

tempdat <- tempdat %>% 
  mutate(date = date(Timestamp),
         AirTC = as.numeric(gsub(x = AirTC, pattern = ",", replacement = "."))) %>%
  group_by(Site, date) %>% 
  summarise(AirTC = mean(AirTC, na.rm = T)) %>% 
  pivot_longer(cols = AirTC, names_to = "Parameter")

regdat <- full_join(precipdat,tempdat)

# Plot regional data if you want

# regdat  %>%
#   filter(month(date)>= 4 & month(date)<=10) %>% 
#   ggplot(aes(x = date, y = value))+
#   geom_line(data = filter(.data = regdat, Parameter == "AirTC"), aes(colour = Site))+
#   geom_col(data = filter(.data = regdat, Parameter == "Rain_Tot"), aes(fill = Site))+
#   geom_line(data = filter(.data = regdat, Parameter == "CuPrecip"), aes(colour = Site))+
#   facet_grid(rows = vars(Parameter),
#              scales = "free")



# -------------------------------------------------------------------------------------------------
# -------------------------------------------------------------------------------------------------
# Read and Tidy the local data

# AF STATION DATA

afraw <- read_csv("Precipitation-Gap-Filling/AF_Station.csv") %>% 
  mutate(value = Value) %>% 
  select(-Value)


# ----------Found an error in Temp data, removing it here  ---------------------------------------


temp.cleaned <- afraw %>% 
  filter(Parameter == "AirTC") %>% 
  filter(Timestamp <= ymd_hms("2020-10-15 19:00:00") | Timestamp >= ymd_hms("2020-10-17 04:00:00")) %>% 
  filter(Timestamp <= ymd_hms("2020-10-20 18:00:00") | Timestamp >= ymd_hms("2020-10-22 00:00:00"))

af.cln <- afraw %>% 
  filter(Parameter != "AirTC")

af.cln <- bind_rows(af.cln, temp.cleaned)

  
# BEAVER LAKE DATA

bvdat <- read_csv("Precipitation-Gap-Filling/BVR_Precip_AirT.csv", skip = 1) %>% 
  slice(-1)

colnames(bvdat) <- c("record","Timestamp","Rain_Tot","AirTC","RH")

bvdat <- bvdat %>% 
  mutate(Timestamp = round_date(mdy_hms(Timestamp),unit = "minute"),
         Site = "BeaverLake") %>% 
  select(-record) %>% 
  filter(Rain_Tot >= 0) %>% 
  pivot_longer(cols = c(Rain_Tot,AirTC,RH), names_to = "Parameter")

# JOIN LOCAL DATA FORMAT FOR FINAL JOIN

localdat <- full_join(af.cln,bvdat) %>%
  mutate(date = date(Timestamp)) %>% 
  filter(Parameter == "AirTC" |
         Parameter == "Rain_Tot" |
         Parameter == "DBTCDT") %>% 
  pivot_wider(names_from = Parameter, values_from = value) %>% 
  group_by(Site,date) %>% 
  summarise(AirTC = mean(AirTC, na.rm = T),
            Rain_Tot = sum(Rain_Tot, na.rm = T),
            DBTCDT = mean(DBTCDT,na.rm = T)) %>% 
  pivot_longer(cols = c(AirTC, Rain_Tot, DBTCDT), names_to = "Parameter")


# -------------------------------------------------------------------------------------------
# -------------------  JOINING ALL DATASETS TOGETHER  ---------------------------------------


dat <- full_join(localdat,regdat)
 

# -------------------------------------------------------------------------------------------
# -------------------  Plot TIMESERIES   ----------------------------------------------------


dat$par.labs <- factor(dat$Parameter,
                           levels = c("AirTC","CuPrecip","Rain_Tot", "DBTCDT"),
                           labels = c(expression(paste("Temp  ","(\u00B0C)")),
                                      expression("CuRain~(mm)"),
                                      expression("Rain~(mm)"),
                                      expression("Snowdepth~(mm)")))



p <- dat %>% 
   # filter(date >= "2020-07-01" & date <= "2021-01-14") 
  filter(Parameter != "CuPrecip") %>% 
  ggplot(aes(x = date, y = value))+
  theme(strip.text.y = element_text( size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8),
        legend.key.size = unit(0.5, "cm"))+
  labs(y = NULL,x = "Date")+
  scale_fill_discrete(name = "Station")+
  scale_colour_discrete(name = "Station")+
  geom_line(data = filter(.data = dat, Parameter == "AirTC"), aes(colour = Site))+
  geom_line(data = filter(.data = dat, Parameter =="DBTCDT"), aes(colour = Site))+
  geom_abline(slope = 0, intercept = 0,linetype = "dashed")+
  geom_col(data = filter(.data = dat, Parameter == "Rain_Tot"), aes(fill = Site))+
  # geom_line(data = filter(.data = dat, Parameter == "CuPrecip"), aes(colour = Site))+
  facet_grid(rows = vars(par.labs),
             scales = "free",
             switch = "both",
             labeller = label_parsed)


ggplotly(p)


```



# Gap Fill the AF Data (Gap from "2020-08-12" to "2020-10-04")




```{r Gap Filling}
# Gap Fill AF Precip Data -------------------------------------------------

# DATE RANGE OF NO AF DATA: raindat$AncientForest[raindat$date>= "2020-08-12" & raindat$date>= "2020-10-04"] = NA
# OVERLAPPING WITH ALL STATions: raindat$date>= "2020-10-04" & raindat$date>= "2020-11-05"]
# LONGEST OVERLAP WITH LONGWORTH & DOMEMTN:  date >= ymd("2020-05-01") & date <= ymd("2020-08-12")

dat %>%
  filter(Parameter == "Rain_Tot") %>% 
  ggplot(aes(x = date,y = value))+
  geom_col(aes(fill = Site))+
  facet_grid(rows = vars(Site))


# Filter to May 2020 - Aug 2020 as this period has long overlap between AF and the MOE sites

raindat <- dat %>% 
  group_by(Site) %>% 
  filter(Parameter == "Rain_Tot") %>%
  filter(date >= ymd("2020-05-01") & date <= ymd("2020-08-12")) %>% 
  filter(Site != "BeaverLake") %>% 
  select(-Parameter, -par.labs) %>%
  mutate(value = RcppRoll::roll_sum(value, n = 3L, align = "right", na.rm = FALSE, fill = numeric(3))) %>% 
  pivot_wider(names_from = Site, values_from = value) %>% 
  pivot_longer(cols = c(-date, -AncientForest, Longworth,DomeMtn), names_to = "Site")

raindat$AncientForest[raindat$date>= ymd("2020-08-12") & raindat$date>= ymd("2020-10-04")] = NA


# Plot the data with fitted line

raindat %>%
  ggplot(aes(x = AncientForest, y = value))+
  geom_point(aes(colour = Site))+
  geom_smooth(method = "lm")+
  facet_grid(cols = vars(Site))
  
# Calculate linear model across dataset


# rainstat <- raindat %>% 
#   nest(-Site) %>% 
#   mutate(fit = map(data, ~ lm(AncientForest ~ value, data = .)),
#          results = map(fit, augment)) %>% 
#   unnest(results,data)
# 
#   
# rainstat  %>% 
#   ggplot(aes(x = date)) +
#   geom_abline(intercept = 0, slope = 1, alpha = .2) +  # Line of perfect fit
#   geom_point(aes(y = .fitted))+
#   geom_point(aes(y = AncientForest), colour = "red")+
#   facet_grid(rows = vars(Site))
# 
# 



# -----------------------------------------------------------------------------------------------------------
# -------------------  Find slope & intercept for Dome & AF relationship ---------------------------------------


rainwide <- dat %>%       
  filter(Parameter == "Rain_Tot") %>%
  filter(date >= ymd("2020-05-01") & date <= ymd("2020-08-12")
  ) %>% 
  select(-Parameter, -par.labs) %>%
  mutate(value = RcppRoll::roll_sum(value, n = 3L, align = "right", na.rm = FALSE, fill = numeric(3))) %>% 
  pivot_wider(names_from = Site, values_from = value) 

rainwide$AncientForest[rainwide$date>= ymd("2020-08-12") & rainwide$date>= ymd("2020-10-04")] = NA


domefit <- rainwide %>% 
  select(AncientForest,DomeMtn)

# -----------------------------------------------------------------------------------------------------------


domefit <- lm(data = domefit, AncientForest~DomeMtn)


summary(domefit)

fitline <- tidy(coefficients(domefit)) %>% 
  pivot_wider(names_from = names, values_from = x, )
colnames(fitline) <- c("b","m")



# -------------------------------------------------------------------------------------------
# -------------------  Calculate the filled values  ---------------------------------------

af.fill <- dat %>% 
  filter(Parameter == "Rain_Tot") %>% 
  filter(date>= "2020-08-12" & date<= "2020-10-04") %>% 
  filter(Site == "DomeMtn") %>% 
  mutate(fil.value = fitline$m*value + fitline$b)

af.fill$fil.value[af.fill$fil.value<0] = 0

af.fill %>% 
  ggplot()+
  geom_point(aes(x = date, y = fil.value))



# -------------------------------------------------------------------------------------------
# -------------------  JOINING Corrected DATASET TOGETHER  ---------------------------------------


af.fill <- af.fill %>% 
  ungroup() %>% 
  mutate(Site = "AncientForest",
         value = fil.value) %>% 
  select(-fil.value)

af.cor <- dat %>%
  filter(Site == "AncientForest") %>% 
  filter(Parameter == "Rain_Tot") %>% 
  filter(date <= "2020-08-11" | date>= "2020-10-05")

af.cor <- bind_rows(af.cor, af.fill)


dat.cor <- dat %>% 
  filter(Site != "AncientForest" | Parameter != "Rain_Tot")

dat.cor <- bind_rows(dat.cor,af.cor)




p <- dat.cor%>% 
  filter(Parameter == "Rain_Tot") %>% 
  ggplot()+
  geom_col(aes(x = date, y = value, colour = Site))

write_csv(dat.cor,"AF_Corr.csv")

ggplotly(p)
  
