---
title: "Cross Correlation"
output:
  html_document:
    df_print: paged
---

```{r Package Loading,include=FALSE}
library(tidyverse)
library(lubridate)
library(plotly)
library(gridExtra)
library(roll)
library(plotly)
library(kableExtra)
library(tseries)
library(tidymodels)

# This script is for cross correlation of rainfall and water level data
```

# 01 Plotting Raw and Rolling Mean data

Some testing was done on length of rolling mean. I applied 6 hrs (24 15min timestamps), 12 hrs (48), and 24 (96). Since early tests of cross correlation indicated that rain preceded water level change by >1 day typically, I have carried out the cross correlation here with 12 hr rolling mean. These top interactive plots just show the difference between the raw and rolling mean data for water level. The final plot is the 12 hr rolling mean (in blue) and raw (faint grey..) rainfall data from Ancient Forest Station.


```{r,echo=FALSE, message=FALSE,warning=FALSE}
rain <- read_csv("Cross-Correlation-Precip-Watertable/AF_Station.csv") %>%
  filter(Parameter == "Rain_Tot") %>% 
  pivot_wider(names_from = "Parameter", values_from = "Value") %>% 
  dplyr::select(-Site) %>% 
  dplyr::rename(timestamp = "Timestamp", rain_mm_tot = "Rain_Tot") 

roll_rain <- rain %>% 
  dplyr::mutate(mn.rain.24 = roll_mean(rain_mm_tot, width = 24),
                mn.rain.48 = roll_mean(rain_mm_tot, width = 48),
                mn.rain.96 = roll_mean(rain_mm_tot, width = 96))


files <- dir(pattern = "*-G.csv", path = "Cross-Correlation-Precip-Watertable/GW-Data") # pull the file names from folder

#files <- "A1-G.csv"


# Correlate and Plot Function ---------------------------------------------


WlBinder <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("Cross-Correlation-Precip-Watertable/GW-Data/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no) %>% 
    dplyr::mutate(mn.val.24 = roll_mean(raw_val, width = 24),
                  mn.val.48 = roll_mean(raw_val, width = 48),
                  mn.val.96 = roll_mean(raw_val, width = 96))
  
  
  # Join the data together and drop all timestamps with NA -----------------
  
  dat <- full_join(roll_rain,wl) %>% 
    drop_na() %>% 
    distinct() %>% 
    filter(date(timestamp) >= "2019-04-01" & date(timestamp) <= "2019-11-01")
  
}

data <- files %>% 
  map(WlBinder) %>% 
  reduce(rbind)

# p24 <- data %>% 
#   ggplot()+
#   geom_line(aes(timestamp, raw_val, colour = site.no), size = 1)+
#   geom_line(aes(timestamp, mn.val.24, group = site.no), linetype="dashed")+
#   labs(title = "6Hr rolling avg")
#   
# ggplotly(p24)

# p48 <- data %>% 
#   ggplot()+
#   geom_line(aes(timestamp, raw_val, colour = site.no), size = 1)+
#   geom_line(aes(timestamp, mn.val.48, group = site.no), linetype="dashed")+
#   labs(title = "12Hr rolling mean Water levels")
#   
# p48
# 
# rain.p48 <- data %>% 
#   ggplot()+
#   geom_col(aes(timestamp,rain_mm_tot))+
#   geom_col(aes(timestamp,mn.rain.48),colour = "#6699CC75")+
#   ylim(c(0,10))+
#   labs(title = "12 hr rolling mean rainfall (blue)")
# rain.p48


```



# 02 Cross Correlation of 6 hr rolling mean 2019 data

## Method

I trimmed the data to the period of 2019-06-01 to 2019-11-01 to avoid snowmelt influence. I then omitted wells B3 (very spiky data) and A3 (sensor drift). I tested for stationarity using Kwiatkowski-Phillips-Schmidt-Shin (KPSS) and Augmented Dickey Fuller (ADF) tests, but found mixed results. Eight sites were found to have stationarity using ADF (p <0.05), but all failed KPSS (p >0.05). I then tried taking first difference of the data and that caused all to become stationary with ADF, and 6 with KPSS.

I'm still unsure of how best to handle detrending in this case, so I've run cross correlation on both the original data (6 hr rolling mean) and the first difference dataset.


### Non-detrended data

```{r 6 hr rolling mean, 01 June to 01 Nov,echo=FALSE, message=FALSE, warning=FALSE}

roll_rain <- rain %>% 
  dplyr::mutate(mn.rain = roll_mean(rain_mm_tot, width = 24)) %>% 
  drop_na()

files <- dir(pattern = "*-G.csv", path = "Cross-Correlation-Precip-Watertable/GW-Data")

binder <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("Cross-Correlation-Precip-Watertable/GW-Data/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no) %>% 
    dplyr::mutate(mn.val = roll_mean(raw_val, width = 24)) %>% 
    drop_na()
}

wldat <- files %>% 
  map(binder) %>% 
  reduce(bind_rows)

# testing for stationarity ---------------------------------------------------

wl.stat <- wldat %>% 
  group_by(site.no) %>% 
  summarise(adf = tidy(adf.test(mn.val))$p.value,
            adf.diff = tidy(adf.test(diff(mn.val,1)))$p.value,
            kpss = tidy(kpss.test(mn.val))$p.value,
            kpss.diff = tidy(kpss.test(diff(mn.val,1)))$p.value)

wl.stat

rain.stat <- roll_rain %>% 
  summarise(adf = tidy(adf.test(mn.rain))$p.value,
            adf.diff = tidy(adf.test(diff(mn.rain,1)))$p.value,
            kpss = tidy(kpss.test(mn.rain))$p.value,
            kpss.diff = tidy(kpss.test(diff(mn.rain,1)))$p.value)

#files <- "A2-G.csv"


# Correlate and Plot Function ---------------------------------------------


CrossCorrelateR <- function(files) {

  sites = substr(files, start = 0, stop = 4)

  # Load in and tidy data ---------------------------------------------------

  wl <- read_csv(paste0("Cross-Correlation-Precip-Watertable/GW-Data/", files)) %>%
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>%
    dplyr::select(timestamp, raw_val,site.no) %>%
    dplyr::mutate(mn.val = roll_mean(raw_val, width = 24))

  wl <- drop_na(wl)

  adf.test(wl$mn.val)
  kpss.test(wl$mn.val)
  # Join the data together and drop all timestamps with NA -----------------

  dat <- full_join(roll_rain,wl) %>%
    drop_na() %>%
    distinct() %>%
    filter(date(timestamp) >= "2019-06-01" & date(timestamp) <= "2019-11-01")

  bccf <-  ccf(dat$mn.rain, dat$mn.val,plot = TRUE, lag = 250, ylab = paste0("ccf ", sites))

  bacfdf <- with(bccf, data.frame(lag, acf))%>%
    rename(crcor = "acf", lagg = "lag") %>%
    mutate(site = sites)

}

data <- files %>%
  map(CrossCorrelateR) %>%
  reduce(rbind) %>%
  filter(site != "B3-G") %>%
  filter(site != "A3-G")




conf.level <- 0.95
ciline <- qnorm((1 + conf.level)/2)/sqrt(14781) # 14781 is the number of data pts in each timeseries

p <- data %>%
  ggplot()+
  geom_line(aes(lagg,crcor,colour = site), size = .9)+
  geom_hline(aes(yintercept = 0)) +
  geom_hline(aes(yintercept = ciline), linetype = 2, color = 'darkblue') +
  geom_hline(aes(yintercept = -ciline), linetype = 2, color = 'darkblue')+
  # geom_segment(mapping = aes(xend = lagg, yend = 0))+
  labs(x = "Lag (15 Min)", y = "Cross-Correlation", title = paste(" 12 hr rollmean Rain to Water Level Correlation 01 Jun - 01 Nov 2019"))

p
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
lag.dat <- data %>%
  dplyr::group_by(site) %>% 
  dplyr::filter(lagg <= 0 & lagg >= -50) %>% 
  dplyr::filter(crcor == max(crcor)) %>%
  dplyr::filter(site != "B3-G" & site != "A3-G") %>% 
  dplyr::mutate(lag.hrs = lagg/4) %>% 
  dplyr::select(site,lagg,lag.hrs,crcor)

kable(lag.dat) %>% 
  kable_material()
```

```{r,echo=FALSE, message=FALSE, warning=FALSE}
lag.summary <- lag.dat %>%
  dplyr::ungroup() %>% 
  dplyr::summarise(mn.lag.hrs = mean(lag.hrs),
                   sd.lag.hrs = sd(lag.hrs))

kable(lag.summary) %>% 
  kable_material()
```

### First difference detrended data

```{r 6 hr rolling mean, 01 June to 01 Nov,echo=FALSE, message=FALSE, warning=FALSE}
########################################################



CrossCorrelateR <- function(files) {

  sites = substr(files, start = 0, stop = 4)
  
  # Load in and tidy data ---------------------------------------------------
  
  wl <- read_csv(paste0("Cross-Correlation-Precip-Watertable/GW-Data/", files)) %>% 
    dplyr::mutate(time = replace_na(time, 00:00:00),
      timestamp = dmy_hms(paste(date,time))) %>% 
    dplyr::select(timestamp, raw_val,site.no) %>% 
    dplyr::mutate(mn.val = roll_mean(raw_val, width = 24))
  
  wl <- drop_na(wl)
  
  adf.test(wl$mn.val)
  kpss.test(wl$mn.val)
  # Join the data together and drop all timestamps with NA -----------------
  
  dat <- full_join(roll_rain,wl) %>% 
    drop_na() %>% 
    distinct() %>% 
    filter(date(timestamp) >= "2019-06-01" & date(timestamp) <= "2019-11-01")
  
  bccf <-  ccf(diff(dat$mn.rain,1), diff(dat$mn.val,1),plot = TRUE, lag = 250, ylab = paste0("diff ccf ", sites))
  
  bacfdf <- with(bccf, data.frame(lag, acf))%>%
    rename(crcor = "acf", lagg = "lag") %>% 
    mutate(site = sites)

}

data <- files %>% 
  map(CrossCorrelateR) %>% 
  reduce(rbind) %>% 
  filter(site != "B3-G") %>% 
  filter(site != "A3-G")


conf.level <- 0.95
ciline <- qnorm((1 + conf.level)/2)/sqrt(14781) # 14781 is the number of data pts in each

p <- data %>% 
  ggplot()+
  geom_line(aes(lagg,crcor,colour = site), size = .9)+ 
  geom_hline(aes(yintercept = 0)) +
  geom_hline(aes(yintercept = ciline), linetype = 2, color = 'darkblue') +
  geom_hline(aes(yintercept = -ciline), linetype = 2, color = 'darkblue')+
  # geom_segment(mapping = aes(xend = lagg, yend = 0))+
  labs(x = "Lag (15 Min)", y = "Cross-Correlation", title = paste(" 12 hr rollmean Rain to Water Level Correlation 01 Jun - 01 Nov 2019"))
  
ggplotly(p)

```


<!-- #  04 Summary Values -->
<!-- ## Peak positive correlation values over 95% confidence (From last plot) -->
<!-- Wells A1-G, B4-G, and C2-G did not have significant correlations over the 2019 period. A1-G water table dynamics are likely different from all other sites due to the site being located in peat as opposed to sediment. The timeseries from B4-G is extremely spiky which appears most likely to be due to sensor noise, though could also be due to an open well head causing artificially rapid recharge. The C transect wells are higher in the flow path, in coarser sediments and within an area with limited ponding. It is not clear why the C2-G correlation does not exceed the 95% CI, however its correlation matches well with others in that transect. -->

```{r,echo=FALSE, message=FALSE, warning=FALSE}
lag.dat <- data %>%
  dplyr::group_by(site) %>% 
  dplyr::filter(lagg <= 0 & lagg >= -50) %>% 
  dplyr::filter(crcor == max(crcor)) %>%
  dplyr::filter(site != "B3-G" & site != "A3-G") %>% 
  dplyr::mutate(lag.hrs = lagg/4) %>% 
  dplyr::select(site,lagg,lag.hrs,crcor)

kable(lag.dat) %>% 
  kable_material()
```


<!-- ## Mean and standard deviation of significant correlations -->
<!-- The time lag between rainfall and water level timeseries correlation is an average of 20.6 hrs, with a standard deviation of 15.6 hrs. This range may be due to a range in infiltration rates (due to range in sediment composition), variability in seal of well head at the ground surface (creating artificially fast recharge pathways, likely at A4-G) or ponding at sites lower in the gradient (rain direct to an above ground water table). -->

```{r,echo=FALSE, message=FALSE, warning=FALSE}
lag.summary <- lag.dat %>%
  dplyr::ungroup() %>% 
  dplyr::summarise(mn.lag.hrs = mean(lag.hrs),
                   sd.lag.hrs = sd(lag.hrs))

kable(lag.summary) %>% 
  kable_material()
```


