---
title: "Theis Well Recovery Process for Estimating Hydraulic Conductivity"
author: "Jeremy Morris"
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,warning = FALSE
)
```

### Read in, organize and plot the data we're using

I've chosen a datum of 5 m below ground surface for hydraulic head measurements so that head values will intuitively make sense (all positive)

## Well recovery B2-G 26 May 2019

This process requires the data to be resampled at a 1-min interval since the 15-min data interval is too coarse. Resampling is done here by spline interpolation.

```{r} 
library(tidyverse)
library(lubridate)
library(zoo)
library(plotly)

files <- dir(path = "Well-Data",pattern = "*.csv")

  # file <- "A2-G_RAW_Appended.csv"

TidyDat <- function(files) {
  
  site <- substr(files,1,4)
  
  dat <- read_csv(paste0("Well-Data/", files)) %>% 
    mutate(site = site,
           Time = replace_na(Time, "00:00:00"),
           datetime = paste(Date,Time),
           datetime = dmy_hms(datetime)) %>% 
    select(site,datetime,Capacitance)
}

dat <- files %>% 
  map(TidyDat) %>% 
  reduce(rbind)

# Calibrate the data
cal <- read_csv("Calibration Values.csv") %>% 
  filter(site == "A2-G" |site == "B2-G" |site == "C2-G")

caldat <- full_join(dat,cal) %>%
  mutate(toctw_mm = m*Capacitance + b,
         dtw_cm = (toctw_mm/10) - toctg_cm,) %>% 
  select(-Capacitance,-serial,-m,-b, -elevation_m,)

rsdat <- caldat %>%
  filter(site == "B2-G") %>% 
  filter(date(datetime) >= "2019-05-26" & date(datetime) <= "2019-05-27")

rsdat <- rsdat %>%
  complete(site, datetime = seq(head(rsdat$datetime,n=1),tail(rsdat$datetime,n=1), by = "1 min")) %>%
  mutate(dtw_cm = na.approx(dtw_cm, na.rm = F),
         head_cm = 500 - dtw_cm,
         time = minutes(datetime)) # this is where resample occurs

p <- rsdat %>% 
  ggplot()+
  geom_point(aes(x = datetime, y =head_cm, colour = site))+
  labs(x = "Date", y = "Hydraulic Head (cm)")

ggplotly(p)

```




### Theis Recovery Equation

Find \(T\), Transmissivity:

\(T = Kb = \frac{2.3Q}{4\pi RDD}log\frac{t}{t'}\)

\(K = \frac{2.3Q}{4\pi\Delta hb}\)

where 
\(K =\) *HydraulicConductivity*,  
\(Q =\) *Pumping Rate*,  
\(RDD =\) *Residual Drawdown*,  
\(t =\) *Time since pumping started*,  
\(t' =\) *Time since pumping stopped*


### Theis Recovery Test Plot


```{r}
ptime <- rsdat %>% 
  mutate(time = minutes(datetime),
    pstart_t = time - as.period(1558878300, "minute"),
           pstop_t = time - as.period(1558880100, "minute"),
         t_t0 = pstart_t/pstop_t,
         logt = log(t_t0),
    RDD = abs(466.1 - head_cm))


plt <- ptime %>% 
  filter(time >= as.period(1558878300, "minute")) %>% 
  ggplot()+
  geom_line(aes(x = t_t0, y = RDD))+
  scale_x_continuous(trans = "log10")+
  labs(title = paste("Theis Recovery Test", ptime$site, date(ptime$datetime)), y = "Residual Drawdown (mm)", x = "T/T' (Unitless)")
ggplotly(plt)


```

## Find difference in \(RDD\) across a complete log cycle

Here I do it between \(0.04<\frac{t}{t'}<0.4\) to get \(RDD = 54.5\).  
I assume aquifer thickness, \(b\), of \(10\ m\) as this is roughly the elevation distance between modern fraser river and the well site elevation.  
I have also assumed an average pumping rate \(Q\) of \(500\ cm^{3} min^{-1} \)

```{r}
Q = 500 #cm3/min
RDD= 54.5 #cm
b = 1000 #cm

K = (2.3*Q)/(4*pi*RDD*b)*(1/60) # conductivity in cm/s
K

```


# Bail Test method
Bail tests work by instantaneously removing a volume from the well

initial head h_{0} is the observed head upon removal of the slug. In our case it will be the lowest point in the raw data prior to the start of recovery.



```{r}


h0 <- 466.05



baildat <- rsdat %>% 
  mutate(h_h0 = head_cm/h0,
         logh = log10(h_h0))

p <- baildat %>% 
  filter(time >= as.period(1558878300, "minute")) %>% 
  ggplot()+
  geom_point(aes(x = datetime, y = h_h0))+
  scale_y_continuous(trans = "log10")
p

ggplotly(p)

r <- 2.54 # well diameter length cm
L <- 20 # well screen length cm
t0 <- "14:45:00"

```

